---
title: "R - Data visualization, linear regression and logistic regression"
author: admin
date: 2021-09-23
image:
  caption: 'Image credit: Pott L.P.'
  focal_point: ""
  placement: 1
  preview_only: false
subtitle: 'Welcome. In this post, I am going to show you how you may create point maps, linear regression and logistic regression utilizing R base tools and tidymodels workflow.'
summary: 'Welcome. In this post, I am going to show you how you may create point maps, linear regression and logistic regression utilizing R base tools and tidymodels workflow.'
  
categories: ["R"]
tags: ["R Markdown", "tidyverse", "maps", "tidymodels"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Data visualization, linear regression and logistic regression
## Crop data collection - ground truth and remote sensing

## Loading the packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(sf)
library(geobr)
```

## URL link of data from data collection and remote sensing on GitHub <https://github.com/luanpott10/Class>

## You will need the raw file <https://raw.githubusercontent.com/luanpott10/Class/main/data_crops.csv>

## Loading the data
```{r}
data <- read.csv("https://raw.githubusercontent.com/luanpott10/Class/main/data_crops.csv")
```

## Plot the data in the map
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE, results='hide'}
cities_RS <- read_municipality(code_muni = "RS", year= 2020)

ggplot()+
  geom_sf(data=cities_RS)+
  geom_point(data=data,aes(x=longitude,y=latitude,fill=class),shape=22,size=2)+
  labs(x= "Longitude", y = "Latitude")+
  scale_fill_manual(values = c("#eded0c", "#49a345"))+
  theme(legend.position = c(0.17, 0.2),
        panel.border = element_rect(color="Black", fill = NA),
        panel.background = element_rect(fill = "#f2f2f2"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.x = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        axis.title.y = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.y = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        legend.title = element_text(family = "serif",
                                    colour = '#000000',size = 12.0),
        legend.text = element_text(family = "serif",
                                   colour = '#000000',size = 12.0),
        legend.background = element_rect(fill="#f2f2f2",
                                  linetype="dashed", 
                                  colour ="#f2f2f2"))
```

## Plot the data variables - continuous x continuous
```{r fig.height=5, fig.width=7}
ggplot(data=data,aes(x=b3_GCVI,y=b4_GCVI))+
  geom_point(aes(fill=class),shape=22,size=2)+
  scale_fill_manual(values = c("#eded0c", "#49a345"))+
  
  stat_smooth(formula = y~x, method="lm", se=FALSE,color="black", linetype='dashed')+

  theme(panel.border = element_rect(color="Black", fill = NA),
        panel.background = element_rect(fill = "#f2f2f2"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.x = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        axis.title.y = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.y = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        legend.title = element_text(family = "serif",
                                    colour = '#000000',size = 12.0),
        legend.text = element_text(family = "serif",
                                   colour = '#000000',size = 12.0),
        legend.background = element_rect(fill="#f2f2f2",
                                  linetype="dashed", 
                                  colour ="#f2f2f2"))
```

## Linear regression
```{r}
lm_fit_x <- lm(b3_GCVI ~ b4_GCVI, data = data)
summary(lm_fit_x)
```

## Linear regression by tidymodels workflow
#### Creating a parsnip specification for a linear regression model
```{r}
lm_model <- linear_reg() |> 
            set_engine('lm') |> 
            set_mode('regression')
```

#### Fitting the model supplying a formula expression and the data
```{r}
lm_fit <- lm_model %>% 
          fit(b3_GCVI ~ b4_GCVI, data = data)
```

#### Summary of the model
```{r}
lm_fit |> 
  pluck("fit") |> 
  summary()

# Also you can use
lm_fit$fit |> summary()
```

#### Parameter estimates of a the lm object
```{r}
tidy(lm_fit)
```

#### Extract the model statistics
```{r}
glance(lm_fit)
```

## Plot the data variables - categorical x continuous
```{r fig.height=5, fig.width=5}
data <- data |> mutate(target = case_when(class == "corn" ~ 0,
                                          class == "soybean" ~ 1))

ggplot(data=data, aes(y=target,x=b4_GCVI))+
  geom_point(aes(fill=as.factor(target)),shape=22,size=2) +
  scale_fill_manual(values = c("#eded0c", "#49a345"))+
  scale_y_continuous(breaks=c(0,1),
                   labels=c("0","1"),
                   limits=c(0,1))+

  stat_smooth(formula = y~x, method="glm", se=FALSE, method.args = list(family=binomial),
              color="black", linetype='dashed')+
  labs(x= "b4_GCVI", y = "Class",fill="Class")+

  theme(legend.position = c(0.17, 0.2),
        panel.border = element_rect(color="Black", fill = NA),
        panel.background = element_rect(fill = "#f2f2f2"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.x = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        axis.title.y = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.y = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        legend.title = element_text(family = "serif",
                                    colour = '#000000',size = 12.0),
        legend.text = element_text(family = "serif",
                                   colour = '#000000',size = 12.0),
        legend.background = element_rect(fill="#f2f2f2",
                                  linetype="dashed", 
                                  colour ="#f2f2f2"))
```

## Logistic regression
```{r}
lg_fit_x <- glm(as.factor(class) ~ b4_GCVI, family="binomial", data=data)
summary(lg_fit_x)
```


## Logistic regression by tidymodels workflow
#### Creating a parsnip specification for a logistic regression model
```{r}
lg_model <- logistic_reg() |> 
        set_engine("glm") |> 
        set_mode("classification")
```

#### Fitting the model supplying a formula expression and the data
```{r}
lg_fit <- lg_model |> 
          fit(as.factor(class) ~ b4_GCVI, data = data)
```

#### Summary of the model
```{r}
lg_fit |> 
  pluck("fit") |> 
  summary()

# Also you can use
lg_fit$fit |> summary()
```

#### Parameter estimates of a the lm object
```{r}
tidy(lg_fit)
```

#### Extract the model statistics
```{r}
glance(lg_fit)
```



