---
title: "R - Cultivated crop maps in Rio Grande do Sul, Brazil"
author: admin
date: 2021-09-22
image:
  caption: 'Image credit: Pott L.P.'
  focal_point: ""
  placement: 1
  preview_only: true
subtitle: 'Welcome. In this post, I am going to show you how you may create maps of cultivated crops in Rio Grande do Sul state, Brazil.'
summary: 'Welcome. In this post, I am going to show you how you may create maps of cultivated crops in Rio Grande do Sul state, Brazil.'
  
categories: ["R"]
tags: ["R Markdown", "tidyverse", "maps"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Generate crop cultivated maps
## Crop plantation area from IBGE 

## Loading the packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(geobr)
library(patchwork)
```

## URL link of data from IBGE on GitHub <https://github.com/luanpott10/Class>

## You will nedd the raw file <https://raw.githubusercontent.com/luanpott10/Class/main/tabela1612_1.csv>

## Load the data and data wrangling
###### The data downloaded from IBGE there are metadata in the 4 first lines that are not intrest for us `skip = 4`.
###### Also, there are info in the last rows, for that we select the 497 municipalities `data[1:497,]`.
###### Furthermore, the rows (cities) without crop planted, IBGE used "-" or "..." instead input 0, then we have used `case_when` function.
```{r message=FALSE, warning=FALSE}
data <- read_csv('https://raw.githubusercontent.com/luanpott10/Class/main/tabela1612_1.csv', skip = 4)
data <- data[1:497,]

colnames(data) <- c("code","city","rice","corn","soybean")

data <- data |> 
  mutate(rice_crop = case_when(rice == "..." | rice == "-" ~ 0,
                                TRUE ~ as.double(rice))) |> 
  mutate(corn_crop = case_when(corn == "..." | corn == "-" ~ 0,
                                TRUE ~ as.double(corn))) |> 
  mutate(soybean_crop = case_when(soybean == "..." | soybean == "-" ~ 0,
                                TRUE ~ as.double(soybean)))
  
data <- data |> select(code,city,rice_crop,corn_crop,soybean_crop)
```

## Dataset of geobr package from municipalities of Rio Grande do Sul state
```{r results='hide', message=FALSE, warning=FALSE}
cities_RS <- read_municipality(code_muni = "RS", year= 2020)
```

## Ggplot of the cities
```{r}
ggplot()+
  geom_sf(data=cities_RS)
```

## Joining the IBGE data and the sf object
```{r message=FALSE, warning=FALSE}
cities_RS$code_muni <- as.character(cities_RS$code_muni)
data_x <- left_join(cities_RS,data,by= c("code_muni"="code"))
```

## Palettes
```{r}
pal_soybean <- c('#252525','#ccece6','#99d8c9','#66c2a4','#41ae76','#238b45','#006d2c','#00441b')
pal_corn <- c('#252525','#f7ffa8','#EFFD5F','#FCE205','#FCD12A','#FFC30B','#F9A602','#c48302')
pal_rice <- c('#252525','#caebfc','#9ecae1','#6baed6','#4292c6','#2171b5','#084594','#082954')
```

## Soybean map
```{r fig.height=5, fig.width=5}
(soybean_map <-
  ggplot()+
    geom_sf(data=data_x,aes(fill=soybean_crop))+
    theme_minimal()+
    scale_fill_gradientn(colours=pal_soybean,
                       limits = c(0,150000),
                       na.value='#252525')+
    labs(x= "Longitude", y = "Latitude", title = "Soybean")+
    guides(fill=guide_colorbar(title="Crop area (ha)",barwidth = 1,barheight = 4,
                             frame.colour = "black"))+
    theme(legend.position = c(0.17, 0.2),
        panel.border = element_rect(color="Black", fill = NA),
        panel.background = element_rect(fill = "#f2f2f2"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.x = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        axis.title.y = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.y = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        legend.title = element_text(family = "serif",
                                    colour = '#000000',size = 12.0),
        legend.text = element_text(family = "serif",
                                   colour = '#000000',size = 12.0),
        legend.background = element_rect(fill="#f2f2f2",
                                  linetype="dashed", 
                                  colour ="#f2f2f2")))
```

## Corn map
```{r fig.height=5, fig.width=5}
(corn_map <-
  ggplot()+
    geom_sf(data=data_x,aes(fill=corn_crop))+
    theme_minimal()+
    scale_fill_gradientn(colours=pal_corn,
                       limits = c(0,15000),
                       na.value='#252525')+
    labs(x= "Longitude", y = "Latitude", title = "Corn")+
    guides(fill=guide_colorbar(title="Crop area (ha)",barwidth = 1,barheight = 4,
                             frame.colour = "black"))+
    theme(legend.position = c(0.17, 0.2),
        panel.border = element_rect(color="Black", fill = NA),
        panel.background = element_rect(fill = "#f2f2f2"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.x = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        axis.title.y = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.y = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        legend.title = element_text(family = "serif",
                                    colour = '#000000',size = 12.0),
        legend.text = element_text(family = "serif",
                                   colour = '#000000',size = 12.0),
        legend.background = element_rect(fill="#f2f2f2",
                                  linetype="dashed", 
                                  colour ="#f2f2f2")))
```

## Rice map
```{r fig.height=5, fig.width=5}
(rice_map <-
  ggplot()+
    geom_sf(data=data_x,aes(fill=rice_crop))+
    theme_minimal()+
    scale_fill_gradientn(colours=pal_rice,
                       limits = c(0,75000),
                       na.value='#252525')+
    labs(x= "Longitude", y = "Latitude", title = "Rice")+
    guides(fill=guide_colorbar(title="Crop area (ha)",barwidth = 1,barheight = 4,
                             frame.colour = "black"))+
    theme(legend.position = c(0.17, 0.2),
        panel.border = element_rect(color="Black", fill = NA),
        panel.background = element_rect(fill = "#f2f2f2"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.x = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        axis.title.y = element_text(family = "serif",
                                    colour = "#000000",size = 22.0),
        axis.text.y = element_text(family = "serif",
                                   colour = "#000000",size = 18.0),
        legend.title = element_text(family = "serif",
                                    colour = '#000000',size = 12.0),
        legend.text = element_text(family = "serif",
                                   colour = '#000000',size = 12.0),
        legend.background = element_rect(fill="#f2f2f2",
                                  linetype="dashed", 
                                  colour ="#f2f2f2")))
```

## Crop maps
```{r fig.height=5, fig.width=15}
soybean_map + corn_map + rice_map
```




