---
title: "R - Data wrangling utilizing the most common functions from tidyverse package collection"
author: admin
date: 2021-10-14
image:
  caption: 'Image credit: Pott L.P.'
  focal_point: ""
  placement: 1
  preview_only: false
subtitle: 'Welcome. In this post, I am going to show you how you may make data wrangling from a dataset utilizing the dplyr, plyr, and tidyr packages from tidyverse package collection.'
summary: 'Welcome. In this post, I am going to show you how you may make data wrangling from a dataset utilizing the dplyr, plyr, and tidyr packages from tidyverse package collection.'
  
categories: ["R"]
tags: ["R Markdown", "data wrangling", "tidyverse"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Data wrangling utilizing the most common functions from tidyverse package collection.

# Loading the package
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

URL link of data from data collection and remote sensing on GitHub <https://github.com/luanpott10/Class>
You will need the raw file <https://raw.githubusercontent.com/luanpott10/Class/main/data_crops.csv>

# Loading the data
```{r}
data <- read.csv("https://raw.githubusercontent.com/luanpott10/Class/main/data_crops.csv")
```

# A rapid view of the data
## Functions that can be used
##### `glimpse()`
##### `head()`
##### `tail()`
##### `summary()`
```{r}
data |> glimpse()

data |> head(5)
data |> tail(5)

data |> summary()
```

# Select
## `select` function is used for selecting columns in a data frame.
## Here I brought you some examples of the `select` function for selecting different columns in different ways.
```{r}
data |> select(1) |> head(6)
data |> select(last_col()) |> head(6)
data |> select(c(class,b0_GCVI,b1_GCVI,b2_GCVI,b3_GCVI,b4_GCVI)) |> head(6)
data |> select(-c(ï..n,latitude,longitude)) |> head(6)

data |> select(starts_with("b")) |> head(6)
data |> select(ends_with("GCVI")) |> head(6)

data |> select(where(is.numeric)) |> head(6)

```

# Arrange
## `arrange` function is used for orders the rows of a data frame by the values of selected columns.
## In some cases we would like to see the ordered rows according a determined column, for that we use `arrange` function.
```{r}
data |> arrange(b4_GCVI) |> head(6)
```

# Relocate
## `relocate` function change column positions.
## Generally in a machine learning model we have the last column as the label, for do that we can use the `relocate` function.
```{r}
data |> relocate(-class) |> head(6)
```

# Filter
## `filter` function is used to subset a data frame, retaining all rows that satisfy your conditions.
## The most common operators that are useful to build the conditions are:
##### `==`, `>`, `>=` 
##### `&`, `|`, `!`, `xor()`
##### `is.na()`
##### `between()`, `near()`
## We also may put more than one filter in the same `filter` function.
```{r}
data |> filter(class == "soybean", b4_GCVI > -4) |> head(6)
```

# Rename
## `rename` changes the names of individual variables.
## Column names in datasets should be short, inuitive and complete. For that many times we need to rename columns for the dataset.
```{r}
data |> rename(n = ï..n) |> head(6)
```

# Summarise
## `summarise` creates a new data frame. 
## The objective of the function is summarise a data frame for an aspect.
## The most common functions used into `summarise` function are:
##### `mean()`, `median()`
##### `sd()`, `IQR()`, `mad()`
##### `min()`, `max()`, `quantile()`
##### `first()`, `last()`, `nth()`
##### `n()`, `n_distinct()`
##### `any()`, `all()`
## We also may put more than one function in `summarise` function.
```{r}
data |> summarise(n = n(), min = min(b4_GCVI), max = max(b4_GCVI), mean = mean(b4_GCVI))
```

# Group by 
## `group_by` takes an existing data frame and converts it into a grouped data frame where operations are performed by group. ungroup() removes grouping.
## Generally the `group_by` function is used before the `summarise` function to generate summaries by group.
```{r}
data |> group_by(class) |> summarise(n = n(), min = min(b4_GCVI), max = max(b4_GCVI), mean = mean(b4_GCVI))
```

# Pull
## `pull` selects a column in a data frame and transforms it into a vector.
## When we are leading with data wrangling we can use `pull` function to extract columns as a vector.
```{r}
data |> pull(b4_GCVI) |> head(6)
```

# Across
## `across` makes it easy to apply the same transformation to multiple columns.
## Generally it is used into `summarise()` and `mutate()` functions.
## In this example we are rounding to 2 decimals the double variables except latitude and longitude.
```{r}
data |> 
  mutate(across(where(is.double) & !c(latitude, longitude), ~ round(.x,2))) |> head(6)
```

# Join
## `join` function joins two data frames together.
## Joining tables, data frames with foreign key, the `by` in the `join` function is the most important for relational databases.
## The types of join are:
### `inner_join` : only rows with matching keys in both x and y;
### `left_join` : all rows in x, adding matching columns from y;
### `right_join` : all rows in y, adding matching columns from x;
### `full_join` : all rows in x with matching columns in y, then the rows of y that don't match x.
## You can se the differences with the below example.
```{r}
data |> select(ï..n,latitude,longitude,class,b0_GCVI) -> data_1
data |> select(ï..n,latitude,longitude,class,b1_GCVI) -> data_2

data_1 <- data_1[1:60,]
data_2 <- data_2[51:100,]

data_x <- inner_join(data_1,data_2, by = c("ï..n","latitude","longitude","class"))
data_x |> glimpse()
data_x <- left_join(data_1,data_2, by = c("ï..n","latitude","longitude","class"))
data_x |> glimpse()
data_x <- right_join(data_1,data_2, by = c("ï..n","latitude","longitude","class"))
data_x |> glimpse()
data_x <- full_join(data_1,data_2, by = c("ï..n","latitude","longitude","class"))
data_x |> glimpse()

```

# Unite
## `unite` function unites the values of two columns into one.
## When we would like to unite two columns to use as a merged column we can use `unite` function.
```{r}
data |> unite ("n_class",ï..n,class,sep="_") -> data_united
data_united |> head(6)
```

# Separate
## `separate` function separates a character column into multiple columns with a regular expression or numeric locations.
## When we would like to separate a column to have two or more columns, we can use the `separate` function
```{r}
data_united |> separate(n_class,c("ï..n","class"),sep="_") |> head(6)
```

# Gather
## `gather` function gathers columns into key-value pairs.
## When we would like to gather columns into two new columns containing a specific column name and the respectively column values we can use the `gather` function.
```{r}
data |> gather(key="Feature",value="Value",b0_GCVI:b4_GCVI) -> data_gathered
data_gathered |> head(6)
```

# Spread
## `spread` function spreads a key-value pair across multiple columns.
## When we need to distribute the pair of key-value columns into a field of cells we can use the `spread` function.
```{r}
data_gathered %>% spread(key = Feature,value = Value) |> head(6)
```

# Final
### Here are the most common functions for data wrangling utilizing tidyverse package.






